\textbf{zklora: Verifiable Multi-Party Inference for Evaluating Private LoRA Weights}

### 2025-01-12 paper/main.tex
more zk details

@@

Notably, although proof-generation times can grow with model size, verification remains comparatively fast (e.g., \texttt{avg\_verify} of about 1--2 seconds even for 70B-scale models). This result demonstrates that \texttt{zklora} allows frequent correctness checks without excessive overhead, which is crucial for real-world “trial runs.” In larger-scale usage scenarios, once trust in the LoRA module is established, \texttt{zklora} may no longer be needed.

%------------------------------------------
% 2. MULTI-PARTY INFERENCE METHOD
%------------------------------------------
\section{Multi-Party Inference: zklora for Verifiable Evaluation}
Our aim is to enable the \emph{Base Model User} to run a \emph{validation experiment} on their local data, incorporating the private LoRA module from the \emph{LoRA Owner}. The Base Model User can measure task performance (e.g., cross-entropy loss, accuracy) to decide whether the LoRA meets performance requirements.

Figure~\ref{fig:toy-mlp-lora} illustrates a simple scenario:
\begin{enumerate}
    \item \textbf{Base Model Forward Pass (Base Model User).}
    The \emph{Base Model User} runs layers without LoRA parameters locally, generating partial activations.

@@

    \item \textbf{LoRA Computation \& Proof Generation (LoRA Owner).}
    The \emph{LoRA Owner} applies the LoRA transformation to the activations and returns the updated activations alongside a \emph{cryptographic proof} of correctness. In our setup, \texttt{zklora} interfaces with \texttt{ezkl} to produce these proofs.

@@

%------------------------------------------
% 3. INTRODUCTION
%------------------------------------------
\section{Introduction}
Large Language Models (LLMs) excel at a variety of tasks but come with significant computational and storage costs \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by introducing a small number of trainable parameters into the model, drastically reducing memory overhead for fine-tuning \cite{ding2022delta}.

In this paper, we present \texttt{zklora}, a library that provides a \emph{verifiable computation} solution for the \emph{Base Model User} to \emph{evaluate} a private LoRA module. By orchestrating a multi-party inference, \texttt{zklora} allows the Base Model User to run the base model locally while sending partial activations to the LoRA Owner. Crucially, the LoRA Owner returns not only updated activations but also a cryptographic \emph{proof} that the correct LoRA transformation was applied. In this way, the Base Model User can confirm the final outputs truly reflect the LoRA module’s capabilities. After verifying and measuring results on a validation dataset, the Base Model User can later choose how to adopt the LoRA at larger scale.

%------------------------------------------
% 4. RELATED WORK
%------------------------------------------
\section{Related Work}

\subsection{Low-Rank Adaptation}
LoRA \cite{hu2021lora} has emerged as a popular way to efficiently fine-tune LLMs by injecting small, low-rank matrices into specific layers. This approach scales well, requiring less GPU memory than full fine-tuning. Various frameworks build on LoRA for GPT-2, GPT-Neo, LLaMA, and others \cite{ding2022delta}. By confining most updated parameters to low-rank adapter matrices, LoRA drastically reduces memory overhead during fine-tuning. As a result, domain-specific LoRA modules can be swapped in or out of the same base model, avoiding the need to retrain or store complete copies of large networks.

\subsection{Incrementally Verifiable Computation}
Valiant~\cite{valiant2008incrementally} introduced the notion of Incrementally Verifiable Computation (IVC), which offers updatable, compact proofs of correctness over multi-step computations. Nova~\cite{kothapalli2022nova} and HyperNova~\cite{kothapalli2024hypernova} refine these ideas, leveraging folding schemes and zero-knowledge properties to enable scaling of verifiable computations in multi-step neural network inference. IVC-based approaches align well with repeated computations needing multiple checkpoints of trust, such as multiple forward passes or partial training loops.

%------------------------------------------
% 5. CONCLUSION
%------------------------------------------
\section{Conclusion}
\texttt{zklora} enables a \emph{Base Model User} to verify the quality of a private LoRA module without revealing the LoRA weights. By pairing multi-party inference with cryptographic proofs, \texttt{zklora} ensures that the user can trust the computed outputs during validation experiments. After verifying performance, the Base Model User may adopt any approach they choose to integrate or deploy the private LoRA more widely, since the module has been proven effective. Our experiments show this approach is feasible across a range of model sizes, with \emph{fast verification times} minimizing the overhead of a verifiable evaluation phase. Future directions include further circuit optimization and exploring multi-owner scenarios, in which multiple private LoRA modules could be composed and verified within the same pipeline.

>>>>

%------------------------------------------
% 4. INTRODUCTION
%------------------------------------------
\section{Introduction}

Large Language Models (LLMs) excel at a variety of tasks but come with significant computational and storage costs \cite{brown2020language, devlin2018bert}. 
Low-Rank Adaptation (LoRA) \cite{hu2021lora} addresses these costs by introducing a small number of trainable parameters into the model, drastically reducing memory overhead for fine-tuning \cite{ding2022delta}.

\texttt{zklora} bridges this gap by orchestrating multi-party inference: the base model user remains in control of the main layers, whereas the LoRA Owner privately applies LoRA transformations. The user obtains a zero-knowledge proof demonstrating that the LoRA transformations are indeed correct, thus verifying the final performance metrics. Once satisfied that the LoRA weights meet their needs, the Base Model User can incorporate them at scale through whatever inference arrangement they prefer.

@@

Notably, although proof-generation times can grow with model size, verification remains comparatively fast (e.g., an average of around 1--2 seconds per proof even for 70B-scale models). This result demonstrates that \texttt{zklora} allows frequent correctness checks without excessive overhead, which is crucial for real-world “trial runs.” In larger-scale usage scenarios, once trust in the LoRA module is established, \texttt{zklora} may no longer be needed.

%------------------------------------------
% 2. MULTI-PARTY INFERENCE METHOD
%------------------------------------------
\section{Multi-Party Inference: zklora for Verifiable Evaluation}

Our aim is to enable the \emph{Base Model User} to run a \emph{validation experiment} on their local data, incorporating the private LoRA module from the \emph{LoRA Owner}. The Base Model User can measure task performance (e.g., cross-entropy loss, accuracy) to decide whether the LoRA meets performance requirements.

Figure~\ref{fig:toy-mlp-lora} illustrates a simple scenario in which:

\begin{enumerate}
    \item \textbf{Base Model Forward Pass (Base Model User).}
    The \emph{Base Model User} runs layers without LoRA parameters locally, generating partial activations.

@@

    \item \textbf{LoRA Computation \& Proof Generation (LoRA Owner).}
    The \emph{LoRA Owner} applies the LoRA transformation to the activations and returns the updated activations alongside a \emph{cryptographic proof} of correctness.

@@

%------------------------------------------
% 3. HOW WE GENERATE THE ZERO-KNOWLEDGE PROOF
%------------------------------------------
\section{Generating the Zero-Knowledge Proof}

A central piece of our methodology is producing a cryptographic proof that the LoRA transformations were correctly computed—without ever disclosing the LoRA parameters themselves. We outline the proof-generation process below, referencing the flow in the accompanying code. For simplicity, we assume the LoRA Owner has already exported the LoRA-modified layer(s) as an ONNX model and that intermediate activations are saved in a JSON file. Each pairing of “model file” and “JSON input data” yields a distinct proof.

\subsection{Preparing Model, Inputs, and Circuit}
\paragraph{Circuit and ONNX File.}
The LoRA Owner provides an ONNX model that includes the relevant LoRA weights. We then compile this model into a zero-knowledge friendly circuit format. This step:
\begin{itemize}
    \item Parses the graph structure from the ONNX file.
    \item Determines how many constraints or “gates” are needed.
    \item Produces a circuit representation that can be used for further proof-generation steps.
\end{itemize}

\paragraph{Input Activations (JSON).}
When the \emph{Base Model User} runs the partial forward pass locally, the intermediate activations for that partial pass are recorded in a JSON file. This JSON effectively captures the inbound tensor(s) that will go through the remote LoRA layers.

\subsection{Generating Setup and Keys}
\paragraph{Settings and Keys.}
To make the cryptographic proof work, the system needs:
\begin{itemize}
    \item \textbf{Settings File:} Summarizes the circuit’s structure (e.g., log rows, public vs. private parameters). 
    \item \textbf{Proving/Verification Keys (PK/VK):} Analogous to a public-private key system, enabling one party to prove a statement and another to verify.
\end{itemize}
We produce these keys from the compiled circuit. If the protocol also requires a structured reference string (SRS), we generate that as well. This SRS encapsulates certain group elements or domain parameters for polynomial commitments, usually saved in a file.

\subsection{Generating the Witness}
\paragraph{Witness Creation.}
With the circuit in place, we feed the JSON data for intermediate activations into the compiled circuit. This step:
\begin{itemize}
    \item Performs the actual forward pass inside the circuit, substituting the LoRA parameters from the ONNX file.
    \item Produces a “witness file,” which is a low-level representation of all wire values in the circuit.
\end{itemize}
In practice, this might be asynchronous or batched, ensuring large inputs can be processed efficiently.

\subsection{Proof Generation}
\paragraph{Prove.}
Once we have:
\begin{itemize}
    \item The circuit representation.
    \item The proving key (PK).
    \item The witness file.
\end{itemize}
we invoke a “prove” routine that executes polynomial commitment schemes and solves the constraint system. This yields a standalone proof file that can be shared with the \emph{Base Model User}.

\paragraph{Result.}
The final proof cryptographically certifies that the LoRA Owner used the declared ONNX model and input JSON to compute the updated activations \emph{exactly} as claimed, with no tampering or hidden changes.

\subsection{Proof Verification}
\paragraph{Checking the Proof.}
The \emph{Base Model User}, upon receiving the proof, uses:
\begin{itemize}
    \item The verification key (VK).
    \item The structured reference string (SRS).
    \item The proof file.
\end{itemize}
A single verification call then confirms or denies the proof’s correctness. If correct, it implies that the LoRA transformations were faithfully applied.

\paragraph{Timing.}
In our benchmarks, this verification step is typically on the order of 1–2 seconds, even for large models, as shown in Table~\ref{tab:results}. The minimal overhead of verification makes \texttt{zklora} feasible for continuous or repeated “trial runs” to confirm LoRA quality.

%------------------------------------------
% 5. RELATED WORK
%------------------------------------------
\section{Related Work}

\subsection{Low-Rank Adaptation}
LoRA \cite{hu2021lora} has emerged as a popular way to efficiently fine-tune LLMs by injecting small, low-rank matrices into specific layers. This approach scales well, requiring less GPU memory than full fine-tuning. Various frameworks build on LoRA for GPT-2, GPT-Neo, LLaMA, and others \cite{ding2022delta}. By confining most updated parameters to low-rank adapter matrices, LoRA drastically reduces memory overhead during fine-tuning.

\subsection{Incrementally Verifiable Computation}
Valiant~\cite{valiant2008incrementally} introduced the notion of Incrementally Verifiable Computation (IVC), which offers updatable, compact proofs of correctness over multi-step computations. Follow-up work \cite{kothapalli2022nova, kothapalli2024hypernova} refines these ideas, leveraging folding schemes and zero-knowledge properties to enable scaling of verifiable computations. While prior research has investigated IVC for inference or specialized tasks, \texttt{zklora} adapts such cryptographic proof techniques for verifying private LoRA transformations in multi-party inference.

%------------------------------------------
% 6. CONCLUSION
%------------------------------------------
\section{Conclusion}

\texttt{zklora} enables a \emph{Base Model User} to verify the quality of a private LoRA module without revealing the LoRA weights. Our benchmarks show that proof generation remains feasible for multiple model sizes, and verification overhead is small enough to allow iterative experiments. This design provides strong assurances of correctness while preserving LoRA privacy. Future directions include optimizing circuit compilation, supporting multi-owner LoRA composition, and investigating broader multi-party computation frameworks that incorporate private data alongside private model parameters.
